Stochastic Blockmodels for Relational Event Data

### 2012-05-27 docs/submission.tex
Minor changes.

Statistical methods for analyzing network data have become increasingly useful for studying  phenomenon ranging from people communicating online to protein interactions \cite{Goldenberg2009}.  One class of statistical models for network data uses latent variables to help model unobserved heterogeneity.  For example, the stochastic blockmodel \cite{Nowicki2001, Kemp, Ishiguro2010} assumes each node in the network belongs to some block (or cluster) and parametrizes the probability of edges between nodes of each block.  % Such approaches are especially useful for large-scale network data where higher-order dependencies can cause models such as exponential random graph models to be too complex to fit.
% TODO: Cite Rodriguez: Modeling dynamics of social networks via h. blockmodels

@@

The above continuous time network models assume all nodes in the network behave according to identical dynamics.  In many cases this is unrealistic.  For example, in a small organization comprised of several teams, each team may uncover different patterns for collaborating via email --  individuals in one group may respond more quickly, while another group may preferentially send to highly active individuals.  %, and in many cases emails between one pair of individuals will not affect emailing behavior between a separate pair of individuals.

>>>>

Statistical methods for analyzing network data have become increasingly useful for studying  phenomenon ranging from people communicating online to protein interactions \cite{Goldenberg2009}.  Stochastic blockmodels \cite{Nowicki2001, Kemp, Ishiguro2010} are a class of statistical models for static network data that employ latent variables to model unobserved heterogeneity by 1) assuming each node in the network belongs to some block (or cluster) and 2) parameterizing the probability of edges between nodes of each block.  % Such approaches are especially useful for large-scale network data where higher-order dependencies can cause models such as exponential random graph models to be too complex to fit.
% TODO: Cite Rodriguez: Modeling dynamics of social networks via h. blockmodels

@@

The above continuous-time network models assume all nodes in the network behave according to identical dynamics.  In many cases this is unrealistic.  For example, in a small organization comprised of several teams, each team may uncover different patterns for collaborating via email --- individuals in one group may respond more quickly, while another group may preferentially send to highly active individuals.

%, and in many cases emails between one pair of individuals will not affect emailing behavior between a separate pair of individuals.
